---
title: "Untitled"
format: html
---

```{r setup}
library(tidyverse)
library(sf)
library(here)
library(qs2)
library(RColorBrewer)

here("src", "dirs.R") %>% source()
here("src", "functions.R") %>% source()
here("src", "display_globals.R") %>% source()
here("src", "mapping_vars.R") %>% source()
```

# Overview

This document imports data from @clawson_mapping_2022 on known and modelled mariculture locations and assigns farms to the species being modelled here. However, there are some challenges to overcome in this process:

- The data from @clawson_mapping_2022 is from 2017 and was based on contemporary aquaculture production data, which has increased significantly since then (and possibly changed in relative species production). New aquaculture locations will not be estimated to accomodate for this - instead, each location will likely produce more than is intended. This is because the aim of this study is to examine change in production/feed demand with climate change, not accurately predict current production.
- There are some gaps acknowledged by @clawson_mapping_2022, particularly surrounding China's production locations.
- There is a possibility that any particular country/FAO fishing region will not have enough farms or farms of a specified type to accomodate all of that country's production. In this case, we will use habitat suitability (temperature range) to place additional farms within that country's EEZ.

# Get data

```{r location-data}
#| code-summary: "Get mariculture locations data"

aqua_locs <- file.path(prepdata_path, "mariculture_locations_FAO_codes.qs") %>% 
    qd_read() %>% 
    rename(ISO3_Code = iso3c)
```

```{r fig-locs}
#| code-summary: "Make sure the mariculture locations data looks right"
#| fig-cap: "Map of mariculture locations, coloured by species group."

p_bigmap_mercator +
  geom_sf(
    data = aqua_locs, 
    aes(color = species_group, geometry = geometry), 
    size = 2
  ) +
  coord_sf() +
  prettyplot() +
  theme(legend.position = "top")
```

```{r production-data}
#| code-summary: "Get prepared production data - already filtered by assigned model names"

FAO_aqua <- file.path(prepdata_path, "FAO_with_models_final_production.csv") %>% 
    read.csv()

# Get FAO fishing region code keys and add to production data
FAO_code_key <- file.path(prepdata_path, "FAO_code_key.csv") %>% 
    read.csv() %>% select(-X)
FAO_aqua <- FAO_aqua %>% 
    left_join(FAO_code_key, by = c("fao_fishing_area" = "NAME_EN"))

# Check that none were left blank
# FAO_aqua %>% 
#     filter(is.na(F_CODE)) %>% 
#     pull(fao_fishing_area) %>% 
#     unique()

# Add groups based on the location data groups
FAO_aqua <- FAO_aqua %>% 
    mutate(
        location_group = case_when(
            model_name %in% c("tuna") ~ "Bluefin tuna",
            model_name %in% c("rainbow_trout", "atlantic_salmon") ~ "Salmonidae fish",
            model_name %in% c("whiteleg_shrimp", "tiger_prawn") ~ "Shrimps and prawns",
            model_name %in% c("mud_crab") ~ "Non-shrimp crustaceans",
            model_name %in% c("japanese_sea_cucumber", "abalone") ~ "Unfed or algae fed bivalve molluscs",
            TRUE ~ "General marine fish",
        ),
    model_name = as.factor(model_name),
    location_group = as.factor(location_group)
    ) %>% 
    select(-c(fao_fishing_area, species_name, environment))

str(FAO_aqua)
```

Production has increased since 2017 (@fig-changes-since-2017). Since the aquaculture mapping data was based on production data, there will be some additional uncertanties based on this change - however the only real implications will be changes in the ratios of species produced. However, from @fig-changes-since-2017, it looks like the changes aren't all that drastic.

```{r fig-changes-since-2017}
#| code-summary: "Show how production of each mapped group has changed in the last 5 years"
#| fig-cap: "Comparing production of each model name from 2017 -- 2023, split by the groupings present in the aquaculture mapping data."

FAO_aqua %>% 
    # I don't know why the years are paired like this in the data but they are - even-numbered years have ~3x the production of odd numbered years
    mutate(period = case_when(
        period %in% 2022:2023 ~ 2023,
        period %in% 2020:2021 ~ 2021,
        period %in% 2018:2019 ~ 2019,
        period %in% 2016:2017 ~ 2017
    )) %>% 
    group_by(period, model_name, location_group) %>% 
    reframe(total_value = sum(value, na.rm = T)) %>% 
    filter(period %in% c(2017:max(FAO_aqua$period))) %>% 
    ggplot(aes(
        x = as.factor(period), 
        y = total_value, 
        fill = model_name
    )) +
    geom_col(colour = "black") +
    facet_wrap(~location_group, scales = "free") +
    scale_fill_manual(
        name = "Model name",
        values = colorRampPalette(brewer.pal(9, "Set1"))(18)
    ) +
    prettyplot() +
    theme(legend.position = "right") +
    labs(
        x = "Year", y = "Production (t)"
    )
```

```{r production-summed}
#| code-summary: "Sum production data from 2017-2023. This will be the production set used going forward."

FAO_aqua_means <- FAO_aqua %>% 
    filter(period >= 2017) %>% 
    group_by(ISO3_Code, country, F_CODE, location_group, model_name) %>% 
    reframe(value = mean(value, na.rm = TRUE)) %>% 
    filter(value != 0)

rm(FAO_aqua)
str(FAO_aqua_means)
```

# Assign locations

For each country and FAO fishing region combination, we follow these steps:
- Get the total number of farms for each `location_group` and the total production of each `model_name`


* If increase ratio > 10, then the existing farm locations are producing too much and new farms need to be added. Habitat suitability data (SST) will be used to assign enough farms to bring the increase ratio down to < 10.
* If increase ratio < 0.1, then the existing farm locations are producing too little and the number of assigned farms needs to be reduced. A subset of the assigned farms will be chosen and the rest will be returned to the "pool" (possibly requiring a `species_group` change).


```{r check-missing-areas}
#| code-summary: "Check which countries/FAO codes are present in each dataset."

all_regions_production <- FAO_aqua_means %>% 
    distinct(ISO3_Code, F_CODE)

all_regions_locations <- aqua_locs %>% 
    distinct(ISO3_Code, F_CODE)

# In production but not locations
in_prod_not_loc <- anti_join(all_regions_production, all_regions_locations, by = c("ISO3_Code", "F_CODE"))

# In locations but not production
in_loc_not_prod <- anti_join(all_regions_locations, all_regions_production, by = c("ISO3_Code", "F_CODE"))

cat(sprintf(
    "In production data but not locations: %d\nIn locations but not production data: %d\n",
    nrow(in_prod_not_loc), nrow(in_loc_not_prod)
))
```

Any farmed locations that don't correspond with any production data will be excluded (they're likely species not included in the `model_names` chosen). Any production values not included in the farmed locations will be included - farmed locations will be generated for them later.

```{r groups-ordered}

# Location groups are handled in order, from most specific -> most general (in terms of the number of model names assigned to them)
groups_order <- c(
    "Bluefin tuna", 
    "Non-shrimp crustaceans", 
    "Salmonidae fish", 
    "Shrimps and prawns", 
    "Unfed or algae fed bivalve molluscs", 
    "General marine fish"
)
```

Each of the following subsections assigns farms for one of these groups, in order. Each subsection gives the following pieces of information:

* `combined_locations_<model>` - a list of the available mapped farms assigned to that model's production globally.
* `regions_without_assignments_<model>` - a list of the regions (countries x FAO regions) which have production data but have no assigned farms. These will need to be assigned using habitat suitability later.
* `farms_excluded_<model>` - the `farm_ID` values of locations that were mapped for a model but which were not selected for this analysis. If needed these can be called upon for future assignments.

## Bluefin tuna

```{r assign-mapped-farms-tuna}
#| code-summary: "Assign mapped farm locations to bluefin tuna."

this_group <- groups_order[1]

all_regions_production <- FAO_aqua_means %>% 
    filter(location_group == this_group) %>% 
    group_by(ISO3_Code, F_CODE) %>% 
    group_split()

all_region_locs <- aqua_locs %>% 
    filter(species_group == this_group) %>% 
    rename(t_per_farm_2017 = tonnes_per_farm) %>% 
    st_drop_geometry() %>% 
    select(-NAME_EN)

plan(
  strategy = "multisession", 
  workers = parallel::detectCores()-2
)

# this_region_prod <- all_regions_production[[4]]

tuna_locations <- future_map_dfr(
    all_regions_production,
    function(this_region_prod) {
        # Info
        ISO3        <- unique(this_region_prod$ISO3_Code)
        F_CODE      <- unique(this_region_prod$F_CODE)

        this_region_locs <- all_region_locs %>% 
            filter(ISO3_Code == ISO3, F_CODE == F_CODE) %>% 
            group_by(species_group) %>% 
            mutate(prop_species_tonnes = sum(t_per_farm_2017, na.rm = T)) %>% 
            ungroup() %>% 
            mutate(prop_species_tonnes = t_per_farm_2017/prop_species_tonnes)

        # If there are any locations to be assigned, they are all assigned to tuna (because it's only one model_name)
        if (nrow(this_region_locs) > 0) {
            production_locations <- this_region_prod %>% 
                left_join(this_region_locs, by = c("ISO3_Code", "country", "F_CODE", "location_group" = "species_group")) %>% 
                mutate(t_per_farm_2023 = prop_species_tonnes * value)
        } else {
            production_locations <- this_region_prod %>% 
                mutate(t_per_farm_2017 = NA, farm_ID = NA, prop_species_tonnes  = NA, t_per_farm_2023 = NA)
        }
        return(production_locations)
    },
    .progress = T,
    .options = furrr_options(seed = TRUE)
)

plan(sequential)
```

```{r check-assignments-tuna}
#| code-summary: "Show how tuna has been assigned to different regions"

# These countries have production values but no locations assigned
tuna_locations %>% 
    filter(is.na(farm_ID)) %>% 
    mutate(number_of_farms = 0) %>% 
    distinct(country, model_name, number_of_farms, value)

# These countries were assigned production - I want to check how many farms were assigned and how much the estimated tonnage has changed from the location to production process. A small increase in production is fine, but a decrease indicates that too many farms were assigned (it's too spread out)
(
    assigned_tuna_summary <- tuna_locations %>% 
        filter(!is.na(farm_ID)) %>% 
        group_by(country, F_CODE, model_name) %>% 
        reframe(
            number_of_farms = n(),
            t_per_farm_2017 = mean(t_per_farm_2017, na.rm = T),
            t_per_farm_2023 = mean(t_per_farm_2023, na.rm = T),
            increase_ratio = t_per_farm_2023/t_per_farm_2017
        ) %>% 
        arrange(-increase_ratio)
)
```

```{r manual-reassignments-tuna}
assigned_tuna_summary <- assigned_tuna_summary %>% 
    filter(increase_ratio < 1)

subs_regions_production <- FAO_aqua_means %>% 
    filter(
        location_group == this_group & 
        country %in% assigned_tuna_summary$country &
        F_CODE  %in% assigned_tuna_summary$F_CODE
    ) %>% 
    group_by(ISO3_Code, F_CODE) %>% 
    group_split()

subs_region_locs <- aqua_locs %>% 
    filter(
        species_group == this_group & 
        country %in% assigned_tuna_summary$country &
        F_CODE  %in% assigned_tuna_summary$F_CODE
    ) %>% 
    rename(t_per_farm_2017 = tonnes_per_farm) %>% 
    st_drop_geometry() %>% 
    select(-NAME_EN)

tuna_locations_2 <- map_dfr(
    subs_regions_production,
    function(this_region_prod) {
        # Info
        ISO3        <- unique(this_region_prod$ISO3_Code)
        F_CODE      <- unique(this_region_prod$F_CODE)
        ir          <- assigned_tuna_summary %>% 
            filter(country == unique(this_region_prod$country) & F_CODE == F_CODE) %>% 
            pull(increase_ratio)

        this_region_locs <- subs_region_locs %>% 
            filter(ISO3_Code == ISO3 & F_CODE == F_CODE) %>% 
            group_by(species_group) %>% 
            mutate(prop_species_tonnes = sum(t_per_farm_2017, na.rm = T)) %>% 
            ungroup() %>% 
            mutate(prop_species_tonnes = t_per_farm_2017/prop_species_tonnes)

        # This is the key part - we're assigning fewer farms than there are
        n_farms     <- as.integer(nrow(this_region_locs) * ir)
        samp_farms  <- sample(this_region_locs$farm_ID, size = n_farms, replace = FALSE)

        production_locations <- this_region_prod %>% 
            left_join(
                filter(this_region_locs, farm_ID %in% samp_farms), 
                by = c("ISO3_Code", "country", "F_CODE", "location_group" = "species_group")
            ) %>% 
            mutate(t_per_farm_2023 = value/n_farms)

        return(production_locations)
    },
    .progress = T
)
```

```{r finish-up-tuna}
#| code-summary: "Collate tuna ssignment data to take forward"

# Which regions need habitat suitability assignment
regions_without_assignments_tuna <- tuna_locations %>% 
    filter(is.na(farm_ID))
tuna_locations <- tuna_locations %>% 
    filter(!is.na(farm_ID))

# Which farms have been excluded and can be returned to the locations pool later (if needed)
farms_excluded_tuna <- setdiff(
    tuna_locations$farm_ID[tuna_locations$ISO3_Code %in% tuna_locations_2$ISO3_Code & tuna_locations$F_CODE %in% tuna_locations_2$F_CODE], 
    tuna_locations_2$farm_ID
)
tuna_locations <- tuna_locations %>% 
    filter(!farm_ID %in% farms_excluded_tuna)

# Update the assigned locations with down-adjusted values
combined_locations_tuna <- rows_update(
  tuna_locations %>% select(-prop_species_tonnes),
  tuna_locations_2 %>% select(-prop_species_tonnes),
  by = c("ISO3_Code", "country", "F_CODE", "location_group", "model_name", "value", "t_per_farm_2017", "farm_ID")
)

combined_locations_tuna %>% 
    group_by(country, F_CODE, model_name) %>% 
    reframe(
        number_of_farms = n(),
        t_per_farm_2017 = mean(t_per_farm_2017, na.rm = T),
        t_per_farm_2023 = mean(t_per_farm_2023, na.rm = T),
        increase_ratio = t_per_farm_2023/t_per_farm_2017
    ) %>% 
    arrange(-increase_ratio)
```

## Non-shrimp crustaceans

```{r assign-mapped-farms-crustaceans}

this_group <- groups_order[2]

all_regions_production <- FAO_aqua_means %>% 
    filter(location_group == this_group) %>% 
    group_by(ISO3_Code, F_CODE) %>% 
    group_split()

all_region_locs <- aqua_locs %>% 
    filter(species_group == this_group) %>% 
    rename(t_per_farm_2017 = tonnes_per_farm) %>% 
    st_drop_geometry() %>% 
    select(-NAME_EN)


```
